Illuminating a Dark Lens : A Type Ia Supernova \\ Magnified by Galaxy Cluster Abell 2744

### 2015-04-01 snTomas.tex
incorporated comments from Julian and Tommaso}

\begin{abstract}
SN \tomas\ is a Type Ia Supernova (\SNIa) discovered at $z=1.31\pm0.01$
behind the galaxy cluster Abell 2744 ($z=0.308$). This SN has a
projected separation from the cluster core of $\sim$40\arcsec, closer
than any other cluster-lensed \SNIa.  As such, it provides the first
opportunity to confront gravitational lens models of galaxy clusters
with the direct measurement of an absolute magnification at the edge
of the strong-lensing region.  We derive a tightly constrained measure
of the peak apparent magnitude, corrected for light curve shape and
exinction.  In a cosmology-independent analysis, we find that \tomas\
is $0.75\pm0.10$ magnitudes brighter than unlensed \SNeIa\ at similar
redshift, implying a lensing magnification of $\mu_{\rm
obs}=2.00\pm0.19$.  We compare this direct measurement of the
magnification against predicted magnifications from 11 lens models,
representing a broad range of methodologies and incorporating a
variety of strong- and weak-lensing constraints.  The models are
collectively fairly accurate, with a naive unweighted mean of
$\mu_{\rm mod}=2.6$.  However, they are systematically biased to
higher values: none of the tested models predict a magnification equal
to or lower than the observed $\mu$, and some models are more than
$5\sigma$ above the measured magnification.  We evaluate possible
causes for this bias, and find that intercomparison of the lens models
rules out several simple possibilities.  A larger sample of
$\sim$10 \SNeIa\ behind any single cluster could provide a robust
evaluation of systematic biases in lens models.  Until such a sample
exists, it is advisable to increase the estimated uncertainty of any
predicted magnification by an amount comparable to the dispersion from
all available lens models.
\end{abstract}

@@

In Section~\ref{sec:DiscoveryAndFollowup} we present the discovery and
follow-up observations of SN \tomas at $z=1.31$, discovered behind the
galaxy cluster Abell 2744.  Sections~\ref{sec:Spectroscopy} and \ref{sec:PhotometricClassification}
describe the spectroscopy and photometry of this SN, leading to a
classification of the object as a normal Type Ia SN.  In
Section~\ref{sec:Magnification} we make a direct measurement of
the magnification of this source due to gravitational lensing, and
compare to predictions from lens models.  Finally,
Section~\ref{sec:Discussion} discusses the tension between our
magnification measurement and the lens models, with implications for
the systematic error budget in magnification estimates for other
high-$z$ lensed sources.

@@

\begin{figure*}
\begin{center}
%\includegraphics[width=\textwidth]{FIG/discovery_image}
\includegraphics[width=\textwidth]{FIG/discovery_image_lowres}
\caption{  \label{fig:DiscoveryImage} 
SN \tomas\ in the Abell 2744 field.  The left panel shows a
UV/Optical/IR color composite image constructed from all available HST
imaging of the Abell 2744 cluster field.  The inset panels at right
show F814W imaging of the immediate vicinity of SN \tomas,
approximately 40\arcsec\ from the center of the cluster. The top panel
shows the template image, combining all data prior to the SN
appearance.  Labeled ellipses mark the nearest galaxies and their
redshift constraints: a photometric redshift for the most likely host
galaxy is marked in blue, and the spectroscopic redshift for a
background galaxy is given in yellow. The bottom panel is constructed
from all HFF F814W imaging taken while the SN was detectable, and
marks the SN location with an arrow.  (Left panel image credit: NASA,
ESA, and J. Lotz, M. Mountain, A. Koekemoer, and the HFF Team (STScI))
}
\end{center}
\end{figure*}

@@

The most probable host galaxy for SN \tomas\ is a faint and diffuse
galaxy immediately to the south-east of the SN location.  With
photometry of the host galaxy collected from the template images, we
fit the spectral energy distribution (SED) using the {\it BPZ} code --
a Bayesian photometric redshift estimator \citep{Benitez:2000}. The
best-fit SED template match is shown in Figure~\ref{fig:HostSED}.
From the {\it BPZ} analysis, we found the host to be most likely an
actively star-forming galaxy at a redshift of $z=1.5\pm0.2$.  Another
nearby bright galaxy to the East of the SN has a spectroscopic
redshift of $z=1.742$, making it most likely a background object.
This redshift was determined from a spectrum taken with the G141 grism
of the HST WFC3-IR camera, collected as part of the Grism Lens
Amplified Survey from Space (GLASS, PI:Treu, PID:13459).

Upon discovery, HST target-of-opportunity observations were triggered
from the FrontierSN program (PI:Rodney, HST-PID:13386), which aims to
discover and follow transient sources in the HFF cluster and parallel
fields. The FrontierSN observations provided WFC3-IR imaging as well
as spectroscopy using the ACS G800L grism, supplementing the
rapid-cadence optical imaging from HST+ACS already being provided by
the HFF program. Difference images for the IR follow-up data were
generated using templates constructed from the HFF WFC3-IR imaging
campaign, which concluded in November, 2013.

@@

The SNID program is designed to estimate the type, redshift, and age
of a SN spectrum through cross-correlation matching with a library of
template spectra.  To account for possible distortions in the broad
shape of the SN pseudo-continuum due to dust or instrumental
calibration effects, SNID divides each SED by a smooth cubic spline
fit. This effectively removes the shape of the SN pseudo-continuum to
leave behind a flat SED superimposed with spectral absorption and
emission features.  It is these features which drive the
cross-correlation fit, so the SNID approach is insensitive to the
overall color of the SED.  We used v2.0 of the SNID template library,
which includes template SEDs covering all Type Ia and Core Collapse
sub-classes, and has recently been updated with corrections and
improvements to the Type Ib/c templates \citep{Liu:2014}.

@@

The {\tt SNID} library can generate an acceptable \CCSN\ match only
when we remove all age and redshift constraints. In this case the best
non-Ia match is the Type Ic SN 1997ef, which delivers {\it rlap}$=6.8$
at $z=0.51$ and age=47.3 rest-frame days past peak.  This is not as
good a fit as the best Type Ia models, is at odds with the host galaxy
redshift prior, and is strongly disfavored by the shape and colors of
the SN light curve (see Section~\ref{sec:PhotometricClassification}).

From the preceding analysis, we conclude that \tomas\ is a Type Ia SN
at $z\approx1.3$.  At this redshift, the absorption at $\sim$8700\AA\
corresponds to the blended \ion{Ca}{2} H\&K features.
This \ion{Ca}{2} absorption is commonly seen in Type Ia SN spectra
near maximum light, although it is also prominent in the spectra of
Type Ib and Ic core collapse SNe (\CCSNe).  The red color of
the \tomas\ SED is qualitatively consistent with a redshift of $z>1$
-- although this information was not used by {\tt SNID} for the
template matching.  As we will see in Section~\ref{sec:PhotometricClassification},
this spectral classification of SN \tomas\ is reinforced by the
photometric information, which also supports classification as a Type
Ia SN at $z=1.3$.

@@

As a check on the spectral classification of SN \tomas\
(Section~\ref{sec:SNID}), we independently classified the SN using a
Bayesian photometric classifier.  We use the
{\tt sncosmo} software
package\footnote{\url{http://sncosmo.github.io/}} to simulate SN light
curves from $z=0.3$ to 2.3 and evaluate the classification
probability using traditional Bayesian model selection \citep[as
in][]{Jones:2013,Rodney:2014,Graur:2014,Rodney:2015}.  In this
analysis we represent normal Type Ia SNe with the SALT2
model \citep{Guy:2010}, and \CCSNe\ with 42 discrete templates (26
Type II and 16 Type Ib/c) drawn from the template library of the
SuperNova Analysis software
package \citep[SNANA,][]{Kessler:2009a}.\footnote{Throughout this work
we use SNANA v10\_35g.}
Likelihoods are defined by comparing the observed fluxes to model
predictions in all passbands where the model is defined.  In practice,
this means we exclude the F435W and F606W bands, which are too blue
for our models at $z>0.85$.

@@

The final photometric classification probability for SN \tomas\ is
$p({\rm Ia}|{\rm {\bf D}})=1.0$, with the classification probability
from all \CCSN\ sub-classes totaling less than $10^{-32}$.  The
best-fit models for each sub-class are shown in
Figure~\ref{fig:photoclass}, which demonstrates that the \CCSN\ models
cannot adequately match the observed photometry. In particular, only
the Type Ia model can simultaneously provide an acceptable fit to the
well-sampled rising light curve in F814W and the F814W-F160W
color near peak.

@@

With the spectroscopic type and redshift securely defined as a normal
Type Ia SN at $z=1.3$, we now turn to fitting the light curve with
Type Ia templates to measure the distance modulus.  Here we use two
independent light curve fitters: the SALT2 model described
above and the MLCS2k2 model \citep{Jha:2007}.

@@

The unlensed sample comprises 18 spectroscopically confirmed Type Ia
SNe in the range $1.14\leq z \leq1.42$ from the GOODS and SCP
surveys,\footnote{GOODS: Great Observatories Origins Deep Survey,
PI:Giavalisco, HST-PID:; SCP: Supernova Cosmology Project,
PI:Perlmutter} which used the HST Advanced Camera for
Surveys \citep{Riess:2006,Suzuki:2012}.  Using the SALT2 and MLCS2k2
fitters as described above, we get distance modulus measures for every
object in this control sample.  We then fit a linear relationship for
distance modulus vs. redshift, and derive a prediction for the
distance modulus of a normal Type Ia SN at the redshift of SN \tomas
(Figure~\ref{fig:MagnificationMeasurement}).
This predicted value is given in
Table~\ref{tab:MeasuredMagnification} under the ``Control'' column.
The difference between the observed distance of SN \tomas\ and this
control sample value is attributed to the magnification from
gravitational lensing:

@@

Before the Frontier Fields observations began, the Space Telescope
Science Institute (STScI) issued a call for lens modeling teams to
generate mass models of all 6 Frontier Field clusters, using a shared
collection of all imaging and spectroscopic data available at the
time.  In response to this opportunity, five teams generated eight
models for Abell 2744.  These models were necessarily quite
preliminary, and they were principally intended to be used for quick
estimation of magnifications for lensed background sources. An
interactive web tool was created by D. Coe and hosted at STScI, to
extract magnification estimates and uncertainties from each model for
any given redshift and position.  In this work we also consider three
additional models that were created later, taking advantage of new
multiply-imaged galaxies discovered in the HFF imaging as well as new
redshifts for lensed background galaxies.

@@

The models shown in Figure~\ref{fig:lensingtest} are separated along
the $y$ direction into two broad classes, with the six {\it
parametric} models on top and the four {\it free-form} models on the
bottom.  Broadly speaking, parametric models use parameterized density
distributions to describe the arrangement of mass within the cluster,
typically associating massive dark matter halos with the positions of
bright cluster member galaxies. Therefore, parametric models rely (to
varying degrees) on the assumption that the observed light from
cluster galaxies is a good tracer of the cluster's dark matter.
Free-form models divide the cluster into a grid, generally using an
adaptive mesh to get better sampling in denser regions. Each grid cell
is assigned a mass, and then the grid spacing and the mass values are
iteratively refined to match the observed lensing constraints.

@@

In spite of the great variety in input data and methodology, these
lens models overall are delivering consistent and fairly accurate
estimates of the magnification.  Furthermore, the scatter amongst
models provides a reasonable estimate of the magnification
uncertainty.  This is largely in agreement with the results
of \citetalias{Patel:2014} and \citet{Nordin:2014}.  The fact that the
model predictions are collectively within $2\sigma$ of the observed
magnification is especially encouraging for these Abell 2744
models. This is a merging cluster with a complex mass distribution,
and most of the models we are evaluating are preliminary products,
generated quickly and without access to the rich data from the
Frontier Fields program.

@@

@@

\begin{enumerate}
\item The close proximity of a cluster member galaxy is leading to a
biased magnification, because that galaxy's M/L is atypical.
\item The models are limited by a scarcity of strong lensing constraints
\item The cluster mass profile exhibits a sudden change outside of the core
strong-lensing region, where we lack multiply-imaged background
sources to constrain it. 
\item \change{The redshift assigned to SN \tomas\ is in error.}
\item \change{SN \tomas\ is not a normal Type Ia SN.}
\end{enumerate}

@@

Increasing the sample of SNe behind well studied strong lensing
clusters would allow this test to be repeated and refined.  With 10 or
100 such objects, it would be possible to see whether the SN \tomas\
$\mu$ discrepancy is simply an outlier, or an indication of a more
pernicious systematic error.  Any cluster that has been vetted by
pencil-beam magnification tests using lensed \SNIa\ will be able to
provide a more reliable measure of the magnifications for very high
redshift objects.  Similarly, a broad sample of SNe like \tomas\ would
help to define the preferred lens modeling methodology by highlighting
any models that consistently perform well in \SNIa\ lensing tests.
The ongoing FrontierSN program will discover and follow any more
highly magnified SNe that appear behind the Frontier Field
clusters. Unfortunately, the HFF survey is not designed with high-z
transient discovery as a primary science goal, so the FrontierSN
effort will likely add no more than 1-3 new lensed \SNIa. Further
imaging of strong-lensing clusters with \HST\ or the James Webb Space
Telescope (JWST) could enable a larger sample to be collected,
especially if the filters and cadence are optimized for detection
of \SNIa\ at $z>1$.  Massive clusters such as Abell 2744 will continue
to be attractive as cosmic telescopes, allowing the next generation of
telescopes to reach the faintest objects in the very early universe.
The puzzling bias revealed by SN \tomas\ supports a concerted effort
to improve these lenses with a larger sample of magnified SNe.

@@

We are p

@@

Financial support for this work was provided by NASA through grants
HST-HF-51312 and HST-GO-13386 from the Space Telescope
Science Institute, which is operated by Associated Universities for
Research in Astronomy, Inc., under NASA contract NAS 5-26555.  Support
for this research at Rutgers University was provided in part by NSF
CAREER award AST-0847157 to SWJ.  The Dark Cosmology Centre is
supported by the Danish National Research Foundation.

@@

>>>>

\begin{abstract}
SN \tomas\ is a Type Ia Supernova (\SNIa) discovered at
$z=1.31\pm0.01$ behind the galaxy cluster Abell 2744 ($z=0.308$). This
SN has a projected separation from the cluster core of
$\sim$40\arcsec, closer to the critical lensing line than any previous
cluster-lensed \SNIa.  As such, it provides the first opportunity to
confront gravitational lens models of galaxy clusters with the direct
measurement of an absolute magnification at the edge of the
strong-lensing region.  We derive a tightly constrained measure of the
peak apparent magnitude, corrected for light curve shape and
exinction.  In a cosmology-independent analysis, we find that \tomas\
is $0.75\pm0.10$ magnitudes brighter than unlensed \SNeIa\ at similar
redshift, implying a lensing magnification of $\mu_{\rm
obs}=2.00\pm0.19$.  We compare this direct measurement of the
magnification against predicted magnifications from 11 lens models,
representing a broad range of methodologies and incorporating a
variety of strong- and weak-lensing constraints. 
The models are collectively fairly accurate, with 3 that are
consistent within the 1$\sigma$ uncertainties.
% a naive unweighted mean of  $\mu_{\rm mod}=2.6$.  
However, the models are systematically biased to higher values: none
of the tested models predict a magnification equal to or lower than
the observed $\mu$, and some models are more than $5\sigma$ above the
measured magnification.  We evaluate possible causes for this bias,
and find that intercomparison of the lens models rules out several
simple possibilities.  A larger sample of $\sim$10 \SNeIa\ behind any
single cluster could provide a robust evaluation of systematic biases
in lens models.  Until such a sample exists, it is advisable
to \change{increase the estimated uncertainty of predicted
magnifications to match the dispersion observed across multiple lens
models.}
\end{abstract}

@@

\change{
One of the key values in observing standard candles behind
gravitational lenses is that {\it all} lensing reconstructions suffer
from the mass-sheet degeneracy \citep{Falco:1985,Bradac:2004}.  This
degeneracy arises because one can introduce into a lens model an
unassociated sheet of mass in front of or behind the lens, without
disturbing the primary observable quantities.  For example, take a
lens model with a given surface mass density $\kappa$, and then
transform the surface mass density to
$\kappa^{\prime}=(1-\lambda)\kappa+\lambda$ for any arbitrary value
$\lambda$.  Both the $\kappa$ and $\kappa^{\prime}$ models will
produce exactly the same values for all positional and shear
constraints from strong and weak lensing.  This is the simplest case
illustrating a positional constraint degeneracy, but there are more
complex versions \citep{Schneider:2014}.  Such mass sheet
degeneracies do not extend to the \citep{absolute} magnification of a
background source's flux and size.  Therefore, an absolute measurement
of magnification from a standard candle or a standard ruler
\citep{Sonnenfeld:2011}  can break this fundamental degeneracy
\citep{Holz:2001}.
}

In Section~\ref{sec:DiscoveryAndFollowup} we present the discovery and
follow-up observations of SN \tomas at $z=1.31$, discovered behind the
galaxy cluster Abell 2744.  Sections~\ref{sec:Spectroscopy}
and \ref{sec:PhotometricClassification} describe the spectroscopy and
photometry of this SN, leading to a classification of the object as a
normal Type Ia SN.  In Section~\ref{sec:LightCurveFitting} we make a
direct measurement of the magnification of this source due to
gravitational lensing, and compare to predictions from lens models.
Finally, Section~\ref{sec:Discussion} discusses the tension between
our magnification measurement and the lens models, with implications
for the systematic error budget in magnification estimates for other
high-$z$ lensed sources.

@@

\begin{figure*}
\begin{center}
%\includegraphics[width=\textwidth]{FIG/discovery_image}
\includegraphics[width=\textwidth]{FIG/discovery_image_lowres}
\caption{  \label{fig:DiscoveryImage} 
SN \tomas\ in the Abell 2744 field.  The left panel shows a
UV/Optical/IR color composite image constructed from all available HST
imaging of the Abell 2744 cluster field.  The inset panels on the right
show F814W imaging of the immediate vicinity of SN \tomas,
approximately 40\arcsec\ from the center of the cluster. The top panel
shows the template image, combining all data prior to the SN
appearance.  Labeled ellipses mark the nearest galaxies and their
redshift constraints: a photometric redshift for the most likely host
galaxy is marked in blue, and the spectroscopic redshift for a
background galaxy is given in yellow. The bottom panel is constructed
from all HFF F814W imaging taken while the SN was detectable, and
marks the SN location with an arrow.  (Left panel image credit: NASA,
ESA, and J. Lotz, M. Mountain, A. Koekemoer, and the HFF Team (STScI))
}
\end{center}
\end{figure*}

@@

The most probable host galaxy for SN \tomas\ is a faint and diffuse
galaxy immediately to the south-east of the SN location.  With
photometry of the host galaxy collected from the template images, we
fit the spectral energy distribution (SED) using the {\it BPZ} code --
a Bayesian photometric redshift estimator \citep{Benitez:2000}. The
best-fit SED template match is shown in Figure~\ref{fig:HostSED}.
From the {\it BPZ} analysis, we found the host to be most likely an
actively star-forming galaxy at a redshift of $z=1.5\pm0.2$.

Another nearby bright galaxy to the East of the SN has a spectroscopic
redshift of $z=1.742$.  This redshift was determined from a spectrum
taken with the G141 grism of the HST WFC3-IR camera, collected as part
of the Grism Lens Amplified Survey from Space (GLASS, PI:Treu,
PID:13459, \citealt{Treu:2015}). \change{As we will see in Sections~\ref{sec:Spectroscopy}
and \ref{sec:PhotometricClassification}, the SN data are inconsistent
with that redshift, meaning that this nearby galaxy is a background
object and therefore has no impact on the SN classification or
luminosity measurement.}

Upon discovery, HST target-of-opportunity observations were triggered
from the FrontierSN program (PI:Rodney, HST-PID:13386), which aims to
discover and follow transient sources in the HFF cluster and parallel
fields. The FrontierSN observations provided WFC3-IR imaging as well
as spectroscopy using the ACS G800L grism, supplementing the
rapid-cadence optical imaging from HST+ACS already being provided by
the HFF program. \change{The last detections in the IR F105W and F140W
bands came from the direct-imaging component of the GLASS program}.
Difference images for the IR follow-up data were generated using
templates constructed from the HFF WFC3-IR imaging campaign, which
concluded in November, 2013.

@@

The SNID program is designed to estimate the type, redshift, and age
of a SN spectrum through cross-correlation matching with a library of
template spectra, \change{using the algorithm of \citet{Tonry:1979}.}
To account for possible distortions in the broad shape of the SN
pseudo-continuum due to dust or instrumental calibration effects, SNID
divides each SED by a smooth cubic spline fit. This effectively
removes the shape of the SN pseudo-continuum to leave behind a flat
SED superimposed with spectral absorption and emission features.  It
is these features which drive the cross-correlation fit, so the SNID
approach is insensitive to the overall color of the SED.  We used v2.0
of the SNID template library, which includes template SEDs covering
all Type Ia and Core Collapse sub-classes, and has recently been
updated with corrections and improvements to the Type Ib/c
templates \citep{Liu:2014}.

@@

Using {\tt SNID} we can find an acceptable \CCSN\ match only
when we remove all age and redshift constraints. In this case the best
non-Ia match is the Type Ic SN 1997ef, which delivers {\it rlap}$=6.8$
at $z=0.51$ and age=47.3 rest-frame days past peak.  This is not as
good a fit as the best Type Ia models, is at odds with the host galaxy
redshift prior, and is strongly disfavored by the shape and colors of
the SN light curve (see Section~\ref{sec:PhotometricClassification}).

From the preceding analysis, we conclude that \tomas\ is a Type Ia SN
at $z\approx1.3$.  At this redshift, the absorption at $\sim$8700\AA\
corresponds to the blended \ion{Ca}{2} H\&K features.
This \ion{Ca}{2} absorption is commonly seen in Type Ia SN spectra
near maximum light, although it is also prominent in the spectra of
Type Ib and Ic core collapse SNe (\CCSNe).  The red color of
the \tomas\ SED is qualitatively consistent with a redshift of $z>1$
-- although this information was not used by {\tt SNID} for the
template matching.  As we will see in Section~\ref{sec:PhotometricClassification},
this spectral classification of SN \tomas\ is reinforced by the
photometric information, which also supports classification as a Type
Ia SN at $z\approx1.3$.

@@

As a check on the spectral classification of SN \tomas\
(Section~\ref{sec:SNID}), we independently classified the SN using a
Bayesian photometric classifier.  We use the
{\tt sncosmo} software
package\footnote{\url{http://sncosmo.github.io/}} to simulate SN light
curves from $z=0.3$ to 2.3 and evaluate the classification
probability using traditional Bayesian model selection \citep[as
in][]{Jones:2013,Rodney:2014,Graur:2014,Rodney:2015a}.  In this
analysis we represent normal Type Ia SNe with the SALT2
model \citep{Guy:2010}, and \CCSNe\ with 42 discrete templates (26
Type II and 16 Type Ib/c) drawn from the template library of the
SuperNova Analysis software
package \citep[SNANA,][]{Kessler:2009a}.\footnote{Throughout this work
we use SNANA v10\_35g.}
Likelihoods are defined by comparing the observed fluxes to model
predictions in all passbands where the model is defined.  In practice,
this means we exclude the F435W and F606W bands, which are too blue
for our models at $z>0.85$.

@@

The final photometric classification probability for SN \tomas\ is
$p({\rm Ia}|{\rm {\bf D}})=1.0$, with the classification probability
from all \CCSN\ sub-classes totaling less than
$10^{-32}$.  \change{Although this Bayesian classification utilizes
the full posterior probability distribution, for illustration we
highlight in Figure~\ref{fig:photoclass} a single best-fit model for
each sub-class.} This demonstrates how the \CCSN\ models fail to
adequately match the observed photometry. In particular, only the Type
Ia model can simultaneously provide an acceptable fit to the
well-sampled rising light curve in F814W and the F814W-F160W color
near peak.

@@

With the spectroscopic type and redshift securely defined as a normal
Type Ia SN at $z=1.31$, we now turn to fitting the light curve with
Type Ia templates to measure the distance modulus.  Here we use two
independent light curve fitters: the SALT2 model described
above and the MLCS2k2 model \citep{Jha:2007}.

@@

The unlensed sample comprises 18 spectroscopically confirmed Type Ia
SNe in the range $1.14\leq z \leq1.42$ from the GOODS and SCP
surveys,\footnote{GOODS: Great Observatories Origins Deep Survey,
PI:Giavalisco, HST-PID:9425,9583; SCP: Supernova Cosmology Project,
PI:Perlmutter} which used the HST Advanced Camera for
Surveys \citep{Riess:2006,Suzuki:2012}.  Using the SALT2 and MLCS2k2
fitters as described above, we get distance modulus measures for every
object in this control sample.  We then fit a linear relationship for
distance modulus vs. redshift, and derive a prediction for the
distance modulus of a normal Type Ia SN at the redshift of SN \tomas
(Figure~\ref{fig:MagnificationMeasurement}).
This predicted value is given in
Table~\ref{tab:MeasuredMagnification} under the ``Control'' column.
The difference between the observed distance of SN \tomas\ and this
control sample value is attributed to the magnification from
gravitational lensing:

@@

Before the Frontier Fields observations began, the Space Telescope
Science Institute (STScI) issued a call for lens modeling teams to
generate mass models of all 6 Frontier Field clusters, using a shared
collection of all imaging and spectroscopic data available at the
time.  In response to this opportunity, five teams generated eight
models for Abell 2744.  \change{These models necessarily relied on
pre-HFF data, and were required to be complete before the HFF program
began, in order to enable the estimation of magnifications for any new
lensed background sources revealed by the HFF imaging.} An interactive
web tool was created by D. Coe and hosted at STScI, to extract
magnification estimates and uncertainties from each model for any
given redshift and position.  In this work we also consider three
additional models that were created later, taking advantage of new
multiply-imaged galaxies discovered in the HFF imaging as well as new
redshifts for lensed background galaxies.

@@

The models shown in Figure~\ref{fig:lensingtest} are separated along
the $y$ direction into two broad classes, with the
six \change{so-called} {\it parametrized} models on top and the four
{\it free-form} models on the bottom.\footnote{\change{Although this
nomenclature is becoming standard in the literature, it is somewhat
misleading, as the pixels or grid cells in free-form models are
effectively parameters as well. Perhaps ``simply-parameterized'' would
be more accurate, though we adopt the more common usage here.}}
Broadly speaking, the parametric models use parameterized density
distributions to describe the arrangement of mass within the cluster,
typically associating massive dark matter halos with the positions of
bright cluster member galaxies. Therefore, parametric models rely (to
varying degrees) on the assumption that the observed light from
cluster galaxies is a good tracer of the cluster's dark matter.
Free-form models divide the cluster into a grid, generally using an
adaptive mesh to get better sampling in denser regions. Each grid cell
is assigned a mass or a potential, and then the grid spacing and the
mass values are iteratively refined to match the observed lensing
constraints.

@@

In spite of the great variety in input data and methodology, these
lens models overall are delivering consistent and fairly accurate
estimates of the magnification.  \change{Four of the tested models
(CATS, GLAFIC, Zitrin-NFW, Merten) report a median value that falls
within the 1$\sigma$ uncertainty range of the measured magnification,
and two of the models (Williams, Merten) report uncertainties that
overlap the central $\mu$ from at least one of the light curve
fitters.}  Furthermore, the scatter amongst
models provides a reasonable estimate of the magnification
uncertainty.

This is largely in agreement with the results
of \citetalias{Patel:2014} and \citet{Nordin:2014}.  The fact that the
model predictions are collectively within $2\sigma$ of the observed
magnification is especially encouraging for these Abell 2744
models. This is a merging cluster with a complex mass distribution,
and most of the models we are evaluating are preliminary products,
generated quickly and without access to the rich data from the
Frontier Fields program.

@@

\subsection{Foreground Dust}
\label{sec:ForegroundDust}

\change{
All SN sight-lines must intersect some amount of foreground dust from
the immediate circumstellar environment, the host galaxy, and the
intergalactic medium (IGM). In the case of SN \tomas\ one might posit
some dust extinction from the intra-cluster medium (ICM) of Abell
2744, although measurements of rich clusters suggest that the ICM has
only a negligible dust
content \citep{Maoz:1995,Stickel:2002,Bai:2007}.  When fitting
the \tomas\ light curve we account for dust by including corrections
that modify the inferred luminosity distance based on the SN color.
If after applying these dust corrections we are still {\it
underestimating} the effect of dust along this sight-line, then the SN
would appear more dim, the inferred distance modulus would be higher,
and the measured magnification would be reduced -- consistent with the
discrepancy we observe.  
}

\change{
One might suppose that a bias could be introduced if we have adopted
incorrect values for the color correction parameter $\beta$ or
extinction law $R_V$ in the SALT2 and MLCS2k2 fits, respectively.  The
appropriate value to use for this color correction and how it affects
inferences about the intrinsic scatter in Type Ia SN luminosities is a
complex question that is beyond the scope of this work \citep[see
e.g.][]{Marriner:2011,Chotard:2011,Kessler:2013,Scolnic:2014a}.
However, we can already rule this out as a solution for the
magnification discrepancy.  Our error on the \tomas\ distance modulus
already includes an uncertainty in the extinction law and a related
error to account for the intrinsic luminosity scatter. These are well
vetted paraemeters, based on observations of $\sim500$ SNe extending
to $z\sim1.5$ \citep{Sullivan:2011}.  Furthermore, changing the color
correction applied to SN \tomas\ would requre the same adjustment to
be applied to the unlensed SNe at similar redshift that make up our
comparison sample, unless one proposes without evidence that
SN \tomas\ is uniquely affected by a peculiar type of dust.  
}

\change{
The traditional color corrections as formalized in SN light curve
fitters are designed to account for a dust component that lies in the
rest frame of the SN.  The inferred luminosity of a SN can also be
affected by the presence of foreground dust with a different redshift
and possibly a different reddening law \citep{Menard:2010b}.  However,
the magnitude of such a bias is insufficient to account for the
observed discrepancy, \citet{Menard:2010a} estimate the opacity of the
universe as $\langle A_{V}\rangle\sim0.03$ mag up to $z=0.5$.  While
this can have a measurable impact on precise cosmological constraints,
it is far less than the 0.285 mag difference between the observed
magnification of \tomas\ and the mean of the model predictions.  
}

\section{Misclassification}
\label{sec:Misclassification}

\textcolor{red}{TODO : discuss 2006bt-like misclassification}

\change{
In Section~\ref{sec:LightCurveFitting} we found that SN \tomas\ is on
the blue end of the normal range of Type Ia SN colors.  With the SALT2
fitter we measured a color parameter $c=-0.115\pm0.025$, and with
MLCS2k2 we found the host galaxy dust extinction to be $A_{\rm
V}=0.014\pm0.037$ magnitudes.  These colors are tightly constrained,
as we are fitting to photometry that covers a rest-frame wavelength
range from $\sim3500-7000$\AA and extends to $\sim$30 days past
maximum brightness.

\textcolor{red}{TODO : discuss 2006bt-like colors}
}

@@

\begin{enumerate}
\item The close proximity of a cluster member galaxy is leading to a
biased magnification, because that galaxy's M/L is atypical.
\item The models are limited by a scarcity of strong lensing constraints
\item The cluster mass profile exhibits a sudden change outside of the core
strong-lensing region, where we lack multiply-imaged background
sources to constrain it. 
\item \change{The redshift assigned to SN \tomas\ is in error.}
\item \change{The color corrections for SN \tomas\ underestimate the foreground dust.}
\item \change{SN \tomas\ is not a normal Type Ia SN.}
\end{enumerate}

@@

\change{
A promising avenue for exploring the origins of systematic biases in
cluster lens models is through the use of simulated data.  One can
start with very deep high-resolution multi-band imaging on an unlensed
field that has a fairly complete spectroscopic redshift catalog, such
as the Hubble Ultra Deep Field.  Then a simulated galaxy cluster is
placed in the field, and the background galaxies are distorted into
arcs and multiple images using a well-defined lensing prescription.
The artificially lensed images can then be distributed to lens
modeling teams who attempt to reconstruct the (known) mass profile of
the simulated cluster.  This exercise has recently been pursued with a
set of synthetic clusters similar to those observed in the HFF program
(Meneghetti et al. in prep).  Preliminary analysis of this simulation
comparison suggests that systematic biases such as we have seen for
SN \tomas\ may be common for sources that lie close to the critical
curve in elongated and highly asymmetric clusters like Abell 2744.
This SN analysis indicates that such simulation efforts will be an
important step for moving toward precision science with cluster-lensed
sources. 
}

Increasing the sample of SNe behind well studied strong lensing
clusters would allow this test to be repeated and refined.  With 10 or
100 such objects, it would be possible to see whether the SN \tomas\
$\mu$ discrepancy is simply an outlier, or an indication of a more
pernicious systematic error.  Any cluster that has been vetted by
pencil-beam magnification tests using lensed \SNIa\ will be able to
provide a more reliable measure of the magnifications for very high
redshift objects.  Similarly, a broad sample of SNe like \tomas\ would
help to define the preferred lens modeling methodology by highlighting
any models that consistently perform well in \SNIa\ lensing tests.
The ongoing FrontierSN program will discover and follow any more
highly magnified SNe that appear behind the Frontier Field
clusters. Unfortunately, the HFF survey is not designed with high-z
transient discovery as a primary science goal, so the FrontierSN
effort will likely add no more than 1-3 new lensed \SNIa. Further
imaging of strong-lensing clusters with \HST\ or the James Webb Space
Telescope (JWST) could enable a larger sample to be collected,
especially if the filters and cadence are optimized for detection
of \SNIa\ at $z>1$.  Massive clusters such as Abell 2744 will continue
to be attractive as cosmic telescopes, allowing the next generation of
telescopes to reach the faintest objects in the very early universe.
The puzzling bias revealed by SN \tomas\ supports a concerted effort
to improve these lenses with \change{further examination of lensing
systematics through simulations and collection of} a larger sample of
magnified SNe.

@@

This work is dedicated to our colleague and friend Tomas Dahlen
(-2014), for whom this supernova has been named in memorium. He is
greatly missed.

@@

Financial support for this work was provided to SAR by NASA through grants
HST-HF-51312 and HST-GO-13386 from the Space Telescope
Science Institute, which is operated by Associated Universities for
Research in Astronomy, Inc., under NASA contract NAS 5-26555.

Support for this research at Rutgers University was provided in part
by NSF CAREER award AST-0847157 to SWJ.  The Dark Cosmology Centre is
supported by the Danish National Research Foundation.

@@

J.M. contributed to this research from the Jet Propulsion Laboratory,
California Institute of Technology, under a contract with NASA and
acknowledges support from NASA Grants HST-GO-13343.05-A and
HST-GO-13386.13-A. The research leading to these results has received
funding from the People Programme (Marie Curie Actions) of the European
Union's Seventh Framework Programme (FP7/2007-­2013) under REA grant
agreement number 627288.
